prolfqua - An R-package for Proteomics Differential Expression Analysis.

### 2022-05-19 inst/MyArticle/paper.Rmd
update formula
Added bioc links

---
title: prolfqua - An R-package for Proteomics Differential Expression Analysis.
author: 
  - Witold E. Wolski $^{\dagger\ddagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - [Swiss Federal Institute of Technology in Zurich](https://ethz.ch) \| [University of Zurich](https://uzh.ch); Winterthurerstrasse 190, CH-8057 Zurich, Switzerland. $^\ddagger$[Swiss Institute of Bioinformatics (SIB)](https://www.sib.swiss/) $^\dagger$Correspondence \texttt{wew@fgcz.ethz.ch}]
  - Paolo Nanni$^{\ast}$
  - Jonas Grossmann$^{\ast\ddagger}$
  - Maria d'Errico$^{\ast\ddagger}$
  - Ralph Schlapbach$^{\ast}$
  - Christian Panse$^{\ast\ddagger}$
output: 
  bookdown::pdf_document2:
    toc: true
bibliography: bibliography.bib
editor_options: 
  chunk_output_type: console
abstract: |
    Protein mass spectrometry is widely used for quantitative proteomics studies, relative protein quantification, and differential expression analysis of proteins. Nevertheless, there is still a need for a flexible and easy-to-use application programmer interface in R that transparently supports a variety of well principled statistical procedures. The package can model simple experimental designs with a single explanatory variable and complex experiments with multiple factors and hypothesis testing.
    The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package integrates, using an R6 application programmer interface, essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation. 
     Furthermore, the package implements benchmark functionality that can help to compare data acquisition, data preprocessing, or data modeling methods using a gold standard dataset. Finally, we show that the implemented methods allow sensitive and specific differential expression analysis. 
     The R package is available through https://github.com/fgcz/prolfqua.
urlcolor: blue
---

@@

When analyzing parallel-group designs using a single explanatory variable, and contrasting groups, we can use all R packages; however, we can use only some of them if we want to model factorial designs or repeated measurements. Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages. We see that packages such as $limma$ and $proDA$ allow us to fit a comprehensive variety of models and test various hypotheses; however, in-depth knowledge of design matrix specification using the R formula interface is required [@law2020guide].

@@

\begin{table}[H]
\centering
\begin{tabular}[H]{ll}
\toprule
Abbreviations & Explaination\\
\midrule
AUC    &Area Under the Curve                               \\
CDF    &Cumulative Distribution Function                   \\
ESI-MS &Electro-Spray-Ionization Mass Spectrometry         \\
$FDP$  &False Discovery Proportion                         \\
$FDR$  &False Discovery Rate                               \\
HPLC   &High-Pressure Liquid Chromatography                \\
LC-MS  &Liquid Chromatography followed by Mass Spectrometry\\
LFQ    &Label-Free Quantification                          \\
LOD    &Limit Of Detection                                 \\
MS     &mass spectrometry                                  \\
OO     &Object-Oriented                                    \\
UML     &       Unified Modeling Language                  \\
\bottomrule
\end{tabular}
\end{table}

@@

# Supplements

>>>>

---
title: prolfqua - An R-package for Proteomics Differential Expression Analysis.
author: 
  - Witold E. Wolski $^{\dagger\ddagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - [Swiss Federal Institute of Technology in Zurich](https://ethz.ch) \| [University of Zurich](https://uzh.ch); Winterthurerstrasse 190, CH-8057 Zurich, Switzerland. $^\ddagger$[Swiss Institute of Bioinformatics (SIB)](https://www.sib.swiss/) $^\dagger$Correspondence \texttt{wew@fgcz.ethz.ch}]
  - Paolo Nanni$^{\ast}$
  - Jonas Grossmann$^{\ast\ddagger}$
  - Maria d'Errico$^{\ast\ddagger}$
  - Ralph Schlapbach$^{\ast}$
  - Christian Panse$^{\ast\ddagger}$
output: 
  bookdown::pdf_document2:
      toc: true
      toc_appendix: false
      toc_bib: false
bibliography: bibliography.bib
editor_options: 
  chunk_output_type: console
abstract: |
    Protein mass spectrometry is widely used for quantitative proteomics studies, relative protein quantification, and differential expression analysis of proteins. Nevertheless, there is still a need for a flexible and easy-to-use application programmer interface in R that transparently supports a variety of well principled statistical procedures. The package can model simple experimental designs with a single explanatory variable and complex experiments with multiple factors and hypothesis testing.
    The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package integrates, using an R6 application programmer interface, essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation. 
     Furthermore, the package implements benchmark functionality that can help to compare data acquisition, data preprocessing, or data modeling methods using a gold standard dataset. Finally, we show that the implemented methods allow sensitive and specific differential expression analysis. 
     The R package is available through https://github.com/fgcz/prolfqua.
urlcolor: blue
---

@@

When analyzing parallel-group designs using a single explanatory variable, and contrasting groups, we can use all R packages; however, we can use only some of them if we want to model factorial designs or repeated measurements. Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages. We see that packages such as `r BiocStyle::Biocpkg("limma")` and `r BiocStyle::Biocpkg("proDA")` allow us to fit a comprehensive variety of models and test various hypotheses; however, in-depth knowledge of design matrix specification using the R formula interface is required [@law2020guide].

@@

\begin{table}[H]
\centering
\begin{tabular}[H]{ll}
\toprule
Abbreviations & Explaination\\
\midrule
AUC    &Area Under the Curve                               \\
CDF    &Cumulative Distribution Function                   \\
ESI-MS &Electro-Spray-Ionization Mass Spectrometry         \\
$FDP$  &False Discovery Proportion                         \\
$FDR$  &False Discovery Rate                               \\
HPLC   &High-Pressure Liquid Chromatography                \\
LC-MS  &Liquid Chromatography followed by Mass Spectrometry\\
LFQ    &Label-Free Quantification                          \\
LOD    &Limit Of Detection                                 \\
MAR    & Missing At Random                                 \\ 
MCAR   & Missing Completely At Random                      \\
MS     &mass spectrometry                                  \\
OO     &Object-Oriented                                    \\
UML     &       Unified Modeling Language                  \\
\bottomrule
\end{tabular}
\end{table}

@@

# Appendix
